Executive Constraints and Civil Conflict Onset

### 2020-01-30 paper.Rmd
First draft presentation of results

---
title: Executive Constraints and Civil Conflict Onset
author: J.L.A. Krusell
date: "`r format(Sys.time(), '%Y-%m-%d')`"
bibliography: library.bib
csl: assets/american-political-science-association.csl
link-citations: true
documentclass: scrreprt
fontsize: 11pt
linestretch: 1.5
header-includes:
    - \usepackage{amsmath}
    - \usepackage{etoolbox}
    - \usepackage{lualatex-math}
    - \usepackage{unicode-math}
    - \appto\Shaded{\singlespacing}
    - \BeforeBeginEnvironment{align*}{\begin{singlespace}\vspace*{-\baselineskip}}
    - \AfterEndEnvironment{align*}{\end{singlespace}\noindent\ignorespaces}
    - \counterwithout{table}{chapter}
    - \counterwithout{figure}{chapter}
output:
    pdf_document:
        fig_caption: true
        keep_tex: true
        latex_engine: lualatex
        toc: true
        toc_depth: 2
        pandoc_args: ["--syntax-definition", "assets/stan.xml",
                      "--highlight-style", "kate",
                      "--filter", "pandoc-crossref"]
    html_document:
        smart: true
        toc: true
abstract: Lorem ipsum
---

[//]: # Setup R environment for entire doc
```{r include = F, results = "hide"}
library(dplyr)
library(extraDistr)
library(ggplot2)
library(gridExtra)
library(magrittr)
library(kableExtra)
library(knitr)
library(tidyr)

@@

[^model]: The model code is written in Stan [@Carpenter2017] and was
    run with the `CmdStan` interface [@Team2019a] using four parallel
    chains at 2000 iterations each with a warmup of 1000. Convergence
    was assessed according to the potential scale reduction factor
    $\hat{R}$ and effective sample size (ESS) [@Vehtari2019] as well as
    standard Stan diagnostics. The full code is available in
    [Appendix A](#appendix-a) and a fully reproducible pipeline
    compromised of replication scripts is available at
    [https://github.com/jsks/ma-thesis](https://github.com/jsks/ma-thesis).

@@

The primary onset dates were determined according to UCDP/PRIO's
episodic definition where a new onset is coded for the same
incompatibility when there is at least a one year pause in the
fighting or drop below the battle death threshold. The primary
rationale is that civil conflicts often go through several distinct
phases of fighting which can make it difficult to accurately code the
end of a conflict. The episodic definition avoids potentially
conflating briefs lulls involving the same rebel groups and leaders
versus the re-emergence of previously long dormant
conflicts. Furthermore, in both scenarios, the theory linking
executive constraints to the incentives for political violence should
still hold (TODO: possible expand here).

That said, because of the potential sensitivity of results to changing
conflict definitions [@Sambanis2004b], all analyses were also run using
only unique onsets.

@@

For the legislature and the judiciary, the two principal government
branches most able to check the power of the executive, there needs to
be both a *capacity* and *willingness* for oversight. Thus, I include
the following legislative variables: *v2lgoppart*, opposition parties
able to exercise oversight; *v2lgfunds*, legislature controls its own
resources; *v2lginvstp*, legislature investigates in practice;
*v2lgqstexp*, legislature questions executive branch officials. For
the judicial component: *v2juhcind*, high court independence;
*v2juncind*, lower court independence; *v2juhccomp*, the degree to
which the government complies with high court decisions; *v2jucomp*,
the degree to which the government complies with lower court
decisions. And for the final dimension, independent investigative
bodies, I include the variable *v2lgovost*, executive oversight by
bodies other than the legislature.

@@

In the preceding equation, $\lambda$ is the variable specific factor
loading, $\gamma$ the population intercepts, and $\psi$ the scale of
the error term. I assume local item independence such that the error
terms, $(\psi_1, \dotsc, \psi_n)$, are uncorrelated (cite Lord &
Novick). Furthermore, for the purposes of identifying the likelihood,
the scale of the latent factor, $\theta$, is fixed to mean $0$ and
standard deviation $1$ while the factor loadings, $(\lambda_1, \dotsc,
\lambda_n)$, are constrained to be positive [@Jackman2009]. The
positive constraints follow theory since all of the manifest variables
exhibit the same scale direction and should be positively correlated
with the latent factor.[^corr] Finally, similar to the regression
coefficients, I select the following weakly informative priors:
\begin{align*}
    \gamma & \sim \mathsf{Normal}(0, 5) \\
    \lambda & \sim \mathsf{LogNormal}(0, 1) \\
    \psi & \sim \mathsf{Weibull}(2, 1)
\end{align*}

[^corr]: In general, the direction of the ordinal answer categories
    for the lower level V-Dem indicators is from low to high, *i.e.*
    $0$ is the "worst" and the highest value the "best." For example,
    for *v2lginvstp*, legislature investigates in practice, $0$
    denotes that it is extremely unlikely that the legislature
    conducts an investigation where the outcome would be unfavorable
    for the sitting executive and the highest category, $4$, denotes
    that it is certain or nearly certain (CITE V-Dem codebook).

@@

In addition to state capacity, I also include a number of standard
variables drawn largely from previous literature. The list of
variables are not meant to provide an all encompassing theory of civil
conflict, but rather roughly capture the opportunity structures for
armed conflict.  This includes a single year lagged measure of GDP per
capita growth from PWT in order to capture short term economic shocks,
which may affect both the strength and legitimacy of the state
[@Bazzi2014]. I also include a measure of mean elevation from the
GROWup dataset [@Girardin2015] to capture terrain advantageous for
conducting an insurgency [@Fearon2003] as well as a measure of
population density derived from population data from PWT and the area
in square kilometers for each country from GROWup. Countries with a
higher population density may require a more extensive and competent
central administration and policing force in order to suppress violent
dissent [@Fearon2003]. There's also the possibility that a high
population density increases the competition for resources. If the
state is identified as the source of a particular scarcity, the risk
of civil conflict may increase. Finally, I include measures of ethnic
fractionalization in the form of the number of political relevant
ethnic groups and the percent of the population facing
discrimination. Where there are significant ethnic cleavages and state
sponsored discrimination, ethnic ties can facilitate collective action
thereby lowering the mobilization costs of rebel groups
[@Cederman2011].

@@

[^peace_yrs]: Regarding the peace years, the variable is calculated as
    the number of years since the last conflict-year, including
    conflict-years from extrasystemic conflicts. For countries left
    censored due to the dataset starting date, I set the start of the
    peace year calculation to 1946, the end of World War II. The
    rationale is that World War II was a significant international
    event that reshaped the world order and even had domestic
    consequences for countries not directly involved. For newly
    independent countries, I use the independence date as the start
    year.

@@

The analysis of the results is split into two sections. First, the
parameter estimates from the executive constraints measurement model
will be presented followed by the parameters pertaining to the
logistic regression as well as a discussion on the overall model fit
to the outcome variable, conflict onset.

@@

```{r fig.cap = "\\label{fig:err} Errors"}
p1 <- ggplot(nonlg_est, aes(se, interval)) + geom_point(col = "#56B4E9", alpha = 0.2, shape = 1)
p2 <- ggplot(nonlg_est, aes(se, diff)) + geom_point(col = "#56B4E9", alpha = 0.2)

@@

```{r fig.width = 7, fig.height = 6, fig.cap = "\\label{fig:theta} Theta"}
theta_summarised <- post_summarise(theta, probs = c(0.10, 0.5, 0.90)) %>%
    rename(codelow = `10%`, median = `50%`, codehigh = `90%`)

@@

To provide a further snapshot of a greater number of countries, figure
\ref{fig:tie} plots the $\theta$ estimates for 2017, the last year in
the dataset, compared to 2007 by country. A full list of all countries
is available in [Appendix C](#appendix-c). Note, the statistical
significance of differences between years for a single country cannot
strictly be judged by overlapping credible intervals (TODO: Schenker
and Gentleman). For example, Romania, despite having overlapping
credible intervals, still has a posterior probability of approximately
$`r romania_p`\%$ that the estimate for 2017 is lower than 2007. In
the case of Chile with even more visual overlap, there is a
$`r chile_p`\%$ posterior probability that the estimate for 2017 is higher
than 2007.

@@

Figure \ref{fig:reg} plots the slope coefficients on the log-odds
scale from the logistic regression linking the latent construct for
executive constraints to the outcome, civil conflict onset. Credible
intervals are summarised the same as previous plots, 80 percent
quantile intervals, and exact values are reported in table
\ref{tab:reg}.

- Log-odds scale
- Standardized

```{r reg, warning = F}
reg <- read_post("posteriors/full_model/reg_posteriors.csv.gz")

@@

probs <- apply(beta, 2, function(x) list(below = mean(x < 0), above = mean(x > 0)))

beta_s <- post_summarise(beta, probs = c(0.1, 0.9)) %>%
    mutate(Estimate = sprintf("[%.3f, %.3f]", signif(`10%`, 3), signif(`90%`, 3))) %>%
    select(-`10%`, -`90%`, -parameter)

rownames(beta_s) <- colnames(beta)

kable(beta_s, "latex", align = "c", booktabs = T, linesep = "", escape = T,
      caption = "Posterior credible intervals for regression slopes") %>%
    kable_styling(latex_options = c("striped", "scale_down"), full_width = F)
```

@@

```{r eval = F}
# Interaction plot
alpha <- extract(reg, "alpha")
theta_seq <- seq(-2.5, 2.5, .01)

@@

f <- function(x, y) {
    with(final.df, alpha +
                   beta[, 1] * x +
                   beta[, 2] * y +
                   beta[, 3] * mean(rgdpepc_gro) +
                   beta[, 4] * mean(pop_density) +
                   beta[, 5] * mean(meanelev) +
                   beta[, 6] * mean(rlvt_groups_count) +
                   beta[, 7] * 1 +
                   beta[, 8] * 1 +
                   beta[, 9] * mean(peace_yrs) +
                   beta[, 10] * 0 +
                   beta[, 11] * x * y)
}

par(mfrow = c(1, 3))
for (y in state) {
    print(y)
    p <- boot::inv.logit(sapply(theta_seq, f, y)) * 100

    m <- apply(p, 2, median)
    pi <- apply(p, 2, quantile, probs = c(0.2, 0.8)) %>% t

    plot(theta_seq, m, type = "l", ylim = c(min(pi[, 1]), max(pi[, 2])),
         xlab = "Executive Constraints", ylab = "Predicted Probability")
    lines(theta_seq, pi[, 1], lty = 2)
    lines(theta_seq, pi[, 2], lty = 2)
}
```

```{r eval = F}
input.df <- readRDS("posteriors/model/final.rds")
p_hat <- readRDS("posteriors/model/summary/p_hat.rds")

plot_dens(p_hat, groups = input.df$lepisode_onset)
```

```{r eval = F}
p <- apply(p_hat, 2, median)
x <- evalmod(scores = p, labels = input.df$lepisode_onset)

plot(x, curvetype = c("ROC", "PRC"))
```

>>>>

---
title: Executive Constraints and Civil Conflict Onset
author: J.L.A. Krusell
date: "`r format(Sys.time(), '%Y-%m-%d')`"
bibliography: library.bib
csl: assets/american-political-science-association.csl
link-citations: true
documentclass: scrreprt
fontsize: 11pt
linestretch: 1.5
header-includes:
    - \usepackage{amsmath}
    - \usepackage{etoolbox}
    - \usepackage{lualatex-math}
    - \usepackage{unicode-math}
    - \appto\Shaded{\singlespacing}
    - \BeforeBeginEnvironment{align*}{\begin{singlespace}\vspace*{-\baselineskip}}
    - \AfterEndEnvironment{align*}{\end{singlespace}\noindent\ignorespaces}
    - \counterwithout{table}{chapter}
    - \counterwithout{figure}{chapter}
output:
    pdf_document:
        fig_caption: true
        keep_tex: true
        latex_engine: lualatex
        toc: true
        toc_depth: 2
        pandoc_args: ["--syntax-definition", "assets/stan.xml",
                      "--highlight-style", "kate"]
    html_document:
        smart: true
        toc: true
abstract: Lorem ipsum
---

[//]: # Setup R environment for entire doc
```{r include = F, results = "hide"}
library(corrplot)
library(dplyr)
library(extraDistr)
library(ggplot2)
library(gridExtra)
library(magrittr)
library(kableExtra)
library(knitr)
library(parallel)
library(precrec)
library(tidyr)

@@

[^model]: The model code is written in Stan [@Carpenter2017] and was
    run with the `CmdStan` interface [@Team2019a] using four parallel
    chains at 2000 iterations each with a warmup of 1000. Convergence
    was assessed according to the potential scale reduction factor
    $\hat{R}$ and effective sample size (ESS) [@Vehtari2019] as well
    as standard Stan diagnostics. The full code is available in
    [Appendix A](#appendix-a) and a fully reproducible pipeline
    compromised of replication scripts is available at
    [https://github.com/jsks/ma-thesis](https://github.com/jsks/ma-thesis). A
    simulation analysis assessing the model and it's ability to
    recover simulated true values is available in [Appendix B](#appendix-b).

[//]: # Load merged dataset for descriptive stats in following sections

```{r results = "hide"}
merged.df <- readRDS("data/merged_data.rds")
```

@@

The primary onset dates were determined according to UCDP/PRIO's
episodic definition where a new onset is coded for the same
incompatibility when there is at least a one year pause in the
fighting or drop below the battle death threshold. The primary
rationale is that civil conflicts often go through several distinct
phases of fighting which can make it difficult to accurately code the
end of a conflict. The episodic definition avoids potentially
conflating briefs lulls involving the same rebel groups and leaders
versus the re-emergence of previously long dormant
conflicts. Furthermore, in both scenarios, the theory linking
executive constraints to the incentives for political violence should
still hold.[^robust]

[^robust]: Due to the potential sensitivity of results to changing
    conflict definitions [@Sambanis2004b], all analyses were also run
    using only unique onsets as well as major episodic onsets with a
    threshold of 1000 battle deaths. The results are available in
    [Appendix C](#appendix-c).

In total there are `r sum(merged.df$episode_onset)` onsets in the
dataset according to the episodic definition affecting
`r filter(merged.df, episode_onset == 1) %$% n_distinct(country_name)`
unique countries out of a total of
`r n_distinct(merged.df$country_name)`. Figure \ref{fig:onsets} plots
the proportion of countries experiencing an ongoing civil conflict and
the proportion of countries experiencing new onsets of conflict over
time.

```{r fig.cap = "\\label{fig:onsets} Proportion of countries with ongoing civil conflicts and new onsets"}

df <- group_by(merged.df, year) %>%
    summarise(`New Onsets` = sum(episode_onset) / n(),
              `Ongoing Conflict` = sum(ongoing) / n()) %>%
    gather(variable, value, -year)

ggplot(df, aes(year, value, colour = variable)) +
    geom_line() +
    theme(legend.title = element_blank()) +
    ylab("Percentage of Countries") +
    xlab("Year")

@@

For the legislature and the judiciary, the two principal government
branches most able to check the power of the executive, there needs to
be both a *capacity* and *willingness* for oversight. Thus, I include
the following legislative variables: *v2lgoppart*, opposition parties
able to exercise oversight; *v2lgfunds*, legislature controls its own
resources; *v2lginvstp*, legislature investigates in practice;
*v2lgqstexp*, legislature questions executive branch officials. For
the judicial component: *v2juhcind*, high court independence;
*v2juncind*, lower court independence; *v2juhccomp*, the degree to
which the government complies with high court decisions; *v2jucomp*,
the degree to which the government complies with lower court
decisions. And for the final dimension, independent investigative
bodies, I include the variable *v2lgotovst*, executive oversight by
bodies other than the legislature.

```{r fa-sum}
manifest <- merged.df %>%
    select(v2lgoppart, v2lgfunds, v2lginvstp, v2lgqstexp, v2juhcind,
           v2juncind, v2juhccomp, v2jucomp, v2lgotovst)

q <- summary_stats(manifest) %>%
    mutate(Variable = paste0("\\textit{", Variable, "}"))

kable(q, "latex", align = "c", booktabs = T, linesep = "", escape = F,
      caption = "Descriptive statistics for manifest variables") %>%
    kable_styling(latex_options = "striped")

Table \ref{tab:fa-sum} provides summary statistics for the manifest
variables. All of the variables are drawn from expert coded surveys
and have been aggregated by V-Dem using IRT models. Thus, they are all
on an unbounded scale with observations roughly falling between -3
and 3. Furthermore, the scale directions are identical and follow the
general direction of the survey answer categories for lower level
V-Dem indicators --- namely, low to high, *i.e.*, 0 is the "worst" and
the highest value the "best." For example, for $v2lginvstp$,
legislature investigates in practice, the answer category 0 denotes
that it is extremely unlikely that the legislature conducts an
investigation where the outcome would be unfavorable for the sitting
executive and the highest category, 4, denotes that it is certain or
nearly certain. As such, on the aggregated, unbounded scale, higher
values in the point estimates denote country-years where the
legislature is more likely to investigate. Finally, figure
\ref{fig:cor} also provides a correlation matrix demonstrating a
fairly high level of correlation across all variables indicating some
commonality in the covariation.

```{r fig.height = 5, fig.width = 7, fig.cap = "\\label{fig:cor} Correlation Heatmap"}
m <- cor(manifest, use = "pairwise.complete.obs") %>%
    signif(3)

corrplot(m, "number", "upper", diag = F, tl.col = "black")

@@

In the preceding equation, $\lambda$ is the variable specific factor
loading, $\gamma$ the population intercepts, and $\psi$ the scale of
the error term. I assume local item independence such that the error
terms, $(\psi_1, \dotsc, \psi_n)$, are uncorrelated (cite Lord &
Novick). Furthermore, for the purposes of identifying the likelihood,
the scale of the latent factor, $\theta$, is fixed to mean $0$ and
standard deviation $1$ while the factor loadings, $(\lambda_1, \dotsc,
\lambda_n)$, are constrained to be positive [@Jackman2009]. The
positive constraints follow theory since, as stated previously, all of
the manifest variables exhibit the same scale direction and should be
positively correlated with the latent factor. Finally, similar to the
regression coefficients, I select the following weakly informative
priors:
\begin{align*}
    \gamma & \sim \mathsf{Normal}(0, 5) \\
    \lambda & \sim \mathsf{LogNormal}(0, 1) \\
    \psi & \sim \mathsf{Weibull}(2, 1)
\end{align*}

@@

In addition to state capacity, I also include a number of standard
variables drawn largely from previous literature. The list of
variables are not meant to provide an all encompassing theory of civil
conflict, but rather roughly capture the opportunity structures for
armed conflict.  This includes a single year lagged measure of GDP per
capita growth from PWT in order to capture short term economic shocks,
which may affect both the strength and legitimacy of the state
[@Bazzi2014]. I also include a measure of mean elevation from the
GROWup dataset [@Girardin2015] to capture terrain advantageous for
conducting an insurgency [@Fearon2003] as well as a measure of
population density derived from population data from PWT and the area
in square kilometers for each country from GROWup. Countries with a
higher population density may require a more extensive and competent
central administration and policing force in order to suppress violent
dissent [@Fearon2003]. There's also the possibility that a high
population density increases the competition for resources. If the
state is identified as the source of a particular scarcity, the risk
of civil conflict may increase. Similarly to GDP per capita, I also
transform mean elevation and population by the natural log. Finally, I
include measures of ethnic fractionalization in the form of the number
of political relevant ethnic groups and the percent of the population
facing discrimination. Where there are significant ethnic cleavages
and state sponsored discrimination, ethnic ties can facilitate
collective action thereby lowering the mobilization costs of rebel
groups [@Cederman2011].

@@

[^peace_yrs]: The peace years variable is calculated as the number of
    years since the last conflict-year, including conflict-years from
    extrasystemic conflicts. For countries left censored due to the
    dataset starting date, I set the start of the peace year
    calculation to 1946, the end of World War II. The rationale is
    that World War II was a significant international event that
    reshaped the world order and even had domestic consequences for
    countries not directly involved. For newly independent countries,
    I use the independence date as the start year.

Table \ref{tab:reg-stats} lists the full descriptive statistics for
all regression covariates.

```{r reg-stats}
df <- merged.df %>%
    select(rgdpepc, rgdpepc_gro, pop_density, meanelev,
           rlvt_groups_count, neighbour_conflict, ongoing, peace_yrs,
           independence)

q <- summary_stats(df)

kable(q, "latex", align = "c", booktabs = T, linesep = "", escape = T,
      caption = "Descriptive statistics for regression covariates") %>%
    kable_styling(latex_options = "striped")

@@

The analysis of the results is split into several sections. First, the
parameter estimates from the executive constraints measurement model
will be presented followed by the parameters pertaining to the
logistic regression and an assessment of model fit. The final section
will be a discussion linking the results back to the outlined theories
as well as an overview of the limitations of the presented model.

@@

```{r fig.cap = "\\label{fig:err} Errors"}
p1 <- ggplot(nonlg_est, aes(se, interval)) +
    geom_point(col = "#56B4E9", alpha = 0.2, shape = 1)

p2 <- ggplot(nonlg_est, aes(se, diff)) +
    geom_point(col = "#56B4E9", alpha = 0.2)

@@

```{r fig.width = 7, fig.height = 5, fig.cap = "\\label{fig:theta} Theta"}
theta_summarised <- post_summarise(theta, probs = c(0.10, 0.5, 0.90)) %>%
    rename(codelow = `10%`, median = `50%`, codehigh = `90%`)

@@

To provide a further snapshot of a greater number of countries, figure
\ref{fig:tie} plots the $\theta$ estimates for 2017, the last year in
the dataset, compared to 2007 by country. A full list of all countries
is available in [Appendix C](#appendix-c). Note, the statistical
significance of differences between years for a single country cannot
strictly be judged by overlapping credible intervals
[@Schenker2001]. For example, Romania, despite having overlapping
credible intervals, still has a posterior probability of approximately
$`r romania_p`\%$ that the estimate for 2017 is lower than 2007. In
the case of Chile with even more visual overlap, there is a
$`r chile_p`\%$ posterior probability that the estimate for 2017 is
higher than 2007.

@@

Figure \ref{fig:reg} plots the slope coefficients on the log-odds
scale from the logistic regression linking the latent construct for
executive constraints to the outcome, civil conflict onset. Credible
intervals are summarised the same as previous plots, 80 percent
quantile intervals, and exact values are reported in table
\ref{tab:reg}  together with the posterior probabilities that each
$\beta$ estimate is greater than zero and less than zero. Recall that
the dependent variables are standardized to have mean $0$ and standard
deviation $1$, with the exception of the ordinal/binary variables
Relevant Group Count, Neighbouring Conflict, Ongoing Conflict, and
Independence.

```{r reg, fig.height = 3, warning = F}
reg <- read_post("posteriors/full_model/reg_posteriors.csv.gz")

@@

# Probabilities that Beta < 0 and Beta > 0
probs <- apply(beta, 2, function(x) {
    list(below = mean(x < 0) %>% round(3) %>% sprintf("%.3f", .),
         above = mean(x > 0) %>% round(3) %>% sprintf("%.3f", .))
})

beta_s <- post_summarise(beta, probs = c(0.1, 0.9)) %>%
    mutate(`Interval Estimate` = sprintf("[%.3f, %.3f]",
                                         signif(`10%`, 3), signif(`90%`, 3)),
           `$Pr(\\beta < 0)$` = lapply(probs, `[[`, "below") %>% unlist,
           `$Pr(\\beta > 0)$` = lapply(probs, `[[`, "above") %>% unlist) %>%
    select(-`10%`, -`90%`, -parameter)

rownames(beta_s) <- colnames(beta) %>% gsub("_", "\\\\_", .)

kable(beta_s, "latex", align = c("c", "r", "r"), booktabs = T,
      linesep = "", escape = F,
      caption = "Posterior credible intervals for regression slopes") %>%
    kable_styling(latex_options = "striped", "scale_down")
```

@@

For the $\beta$ parameter for executive constraints, this means that
there is an $80\%$ posterior probability that an increase of one
standard deviation decreases the log-odds of conflict onset between
$`r beta_s[1, 1]`$. A one standard deviation increase in this context
corresponds roughly to the difference between the level of executive
constraints for Zambia in 1967 under the administration of Kenneth
Kaunda right before the banning of opposition parties following the
1968 election with a score of approximately $0$ and Greece in 1976
shortly after the end of military rule and establishment of a
democratic republic and with a score of approximately $1$. The
probability that the $\beta$ estimate for executive constraints is
below $0$ is $0.984$.

```{r results = "hide"}
merged.df <- readRDS("data/merged_data.rds")

gdp_nigeria <- filter(merged.df, country_name == "Nigeria", year == 2014) %$%
    rgdpepc %>%
    prettyNum(big.mark = ",")

gdp_czech <- filter(merged.df, country_name == "Czech Republic", year == 1992) %$%
    rgdpepc %>%
    prettyNum(big.mark = ",")

gdpgro_korea <- filter(merged.df, country_name == "South Korea", year == 1990) %$%
    rgdpepc_gro %>%
    multiply_by(100) %>%
    signif(3)

gdpgro_usa <- filter(merged.df, country_name == "United States of America", year == 1992) %$%
    rgdpepc_gro %>%
    multiply_by(100) %>%
    signif(3)

lth <- filter(merged.df, country_name == "Lithuania", year == 2012) %>%
    select(pop, area_sqkm) %>%
    mutate(area_sqlm = signif(area_sqkm, 3))

india <- filter(merged.df, country_name == "India", year == 1978) %>%
    select(pop, area_sqkm, meanelev) %>%
    mutate(area_sqkm = signif(area_sqkm, 3),
           meanelev = signif(area_sqkm, 3))

afghan <- filter(merged.df, country_name == "Afghanistan", year == 2004) %>%
    select(meanelev) %>%
    mutate(meanelev = signif(meanelev, 3))

The slopes for the remaining covariates generally following the
intuitive expectation of their relationship to civil conflict onset
with high probabilities of being either greater than or less than
zero. For example, GDP per capita is negative to which a one unit
increase corresponds approximately to Nigeria in 2014 with a value of
$`r gdp_nigeria`$ and Czech Republic in 1992 after the fall of
communism with a value of $`r gdp_czech`$. In other words, an increase
in GDP per capita on the logarithmic scale decreases the log-odds of
conflict onset. Similarly, GDP per capita growth is negatively
associated with conflict onset with an increase from 0 to 1 on the
standardized scale matching the level of United States in 1992,
$`r gdpgro_usa`$ percent, and South Korea in 1990, $`r gdpgro_korea`$ percent.
Population density, mean elevation, number of politically relevant
ethnic groups, presence of neighbouring conflict, and the dummy
variable for independence year all increase the risk of civil conflict
onset. A one unit increase in population density is similar to going
from Lithuania in 2012 with a population of $`r lth$pop`$ million and
an area of $`r lth$area_sqkm`$ square kilometers to India in 1978 with
a population of $`r india$pop`$ million and an area of
$`r india$area_sqkm`$ square kilometers. A one unit increase for mean
elevation roughly matches the difference between India,
$`r india$meanelev`$ meters, and Afghanistan, $`r afghan$meanelev`$
meters. Finally, the number of peace years since the last civil
conflict and the dummy variable for ongoing civil conflict decrease
the risk of conflict onset. This latter result, though surprising, is
a result of conditioning on the other covariates.

```{r results = "hide"}
gdp_eth <- filter(merged.df, country_name == "Ethiopia", year == 2012) %$%
    rgdpepc %>%
    prettyNum(big.mark = ",")

gdp_india <- filter(merged.df, country_name == "India", year == 2014) %$%
    rgdpepc %>%
    prettyNum(big.mark = ",")

gdp_nz <- filter(merged.df, country_name == "New Zealand", year == 2002) %$%
    rgdpepc %>%
    prettyNum(big.mark = ",")
```

To interpret the results for the interaction term between the slopes
for executive constraints and GDP per Capita I calculate the marginal
effect of $\theta$, executive constraints, on the predicted
probability of civil conflict onset for different levels of GDP per
capita while holding constant the other covariates.  In this case, I
took the mean of GDP per capita growth, population density, mean
elevation, politically relevant ethnic group count, and peace years
while setting the dummy indicators for independence, ongoing conflict,
and neighbouring conflict and the random intercepts for year and
country to 0. The levels of GDP per capita were selected based on the
10\%, 50\%, and 90\% quantiles of the standardized log transformed
values. These levels roughly correspond to the GDP per capita of
Ethiopia in 2012 at $`r gdp_eth`$, India in 2014 at $`r gdp_india`$,
and New Zealand in 2002 at $`r gdp_nz`$. In the dataset there are $75$
conflict episode onsets below the first level, $198$ between the first
and second, $73$ between the second and third, and $6$ above the
third.

```{r fig.cap = "\\label{fig:int} Interaction between executive constraints and GDP per capita with 80\\% CIs"}
# Interaction plot
alpha <- extract(reg, "alpha")
theta_seq <- seq(-2.5, 2.5, .01)

@@

f <- function(x, y) {
    with(final.df, alpha +
                   beta[, 1] * x +
                   beta[, 2] * y +
                   beta[, 3] * mean(rgdpepc_gro) +
                   beta[, 4] * mean(pop_density) +
                   beta[, 5] * mean(meanelev) +
                   beta[, 6] * mean(rlvt_groups_count) +
                   beta[, 7] * 0 +
                   beta[, 8] * 0 +
                   beta[, 9] * mean(peace_yrs) +
                   beta[, 10] * 0 +
                   beta[, 11] * x * y)
}

df <- lapply(state, function(y) {
    mclapply(theta_seq, function(theta) {
        p_hat <- boot::inv.logit(f(theta, y)) * 100
        post_summarise(p_hat, probs = c(0.1, 0.5, 0.9)) %>%
            mutate(theta = theta,
                   y = y)
    }, mc.cores = 4) %>% bind_rows
}) %>% bind_rows

df %<>%
    rename(codelow = `10%`,
           median = `50%`,
           codehigh = `90%`) %>%
    mutate(y = case_when(y == state[1] ~ "Low GDP per Capita",
                         y == state[2] ~ "Median GDP per Capita",
                         y == state[3] ~ "High GDP per Capita") %>%
               factor(levels = c("Low GDP per Capita",
                                 "Median GDP per Capita",
                                 "High GDP per Capita")))

ggplot(df, aes(theta, median)) +
    geom_line() +
    geom_ribbon(aes(ymin = codelow, ymax = codehigh), alpha = 0.2) +
    facet_grid(. ~ y) +
    xlab("Executive Constraints") +
    ylab("Predicted Probability")
```

From the results in figure \ref{fig:int} it is clear that the overall
effect of executive constraints when interacted with GDP per capita on
the predicted probability of conflict onset is negative. For countries
with a relatively high level of GDP per capita, going from the worst
to best in executive constraints decreases the predicted probability
of conflict onset from [1.5, 4.5] down to below 0.5. This effect
diminishes, however, for lower values of GDP per capita to the point
that executive constraints are practically indeterminate to the risk
of conflict onset for low-income country-years.

### Assessing Model Fit

Figure \ref{fig:phat} plots the densities of the first 50 posterior
draws for the predicted probabilities of conflict onset for two
groups, country-years that actually experienced an onset of a civil
conflict and country-years that did not. On average, the predicted
probabilities for the actual onset country-years are higher albeit
still extremely low. This can be seen in the summary statistics
reported in table \ref{tab:psum} for the posterior means of the
predicted probabilities. For onset country-years the interquartile
interval is only 0.0485 to 0.209 with a max of 0.556. To better
compare the predicted probabilities to the actual data, table
\ref{tab:confusion} presents a confusion matrix with a 0.2 cutoff,
which corresponds to the third quartile in the summary table.
`r signif((93 / (259 + 93)) * 100, 3)` percent of onsets and
`r signif((7902 / (7902 + 172)) * 100, 3)` percent of non-onsets are
correctly predicted. The false positive rate is
`r signif((172 / (7902 + 172)) * 100, 3)` percent and the false
negative rate is `r signif((259 / (93 + 259)) * 100, 3)` percent.

```{r fig.height = 2, fig.cap = "\\label{fig:phat} Density plot for predicted probabilities"}
p_hat <- extract(reg, "p_hat")

plot_dens(p_hat, groups = final.df$lepisode_onset) + xlim(c(0, 0.4))
```

```{r psum, warning = F}
p_mean <- colMeans(p_hat)

df <- bind_rows(summary(p_mean[final.df$lepisode_onset == 1]),
                summary(p_mean[final.df$lepisode_onset == 0])) %>%
    mutate_all(~signif(., 3))

rownames(df) <- c("Conflict Onset", "No Onset")

kable(df, "latex", align = "c", booktabs = T, linesep = "", escape = F,
      caption = "Summarized posterior predicted probabilities")
```

```{r confusion}
m <- table(p_mean > .2, final.df$lepisode_onset)

rownames(m) <- c("Predicted: No Onset", "Predicted: Onset")
colnames(m) <- c("Actual: No Onset", "Actual: Onset")

kable(m, "latex", align = "c", booktabs = T, linesep = "", escape = F,
      caption = "Confusion matrix with a 0.2 cutoff")
```

```{r results = "hide"}
x <- evalmod(scores = p_mean, labels = final.df$lepisode_onset)

auc <- attr(x$roc[[1]], "auc") %>% signif(3)
auc_pr <- attr(x$prcs[[1]], "auc") %>% signif(3)
```

Figure \ref{fig:roc} plots a receiver operating characteristic (ROC)
curve depicting the tradeoff between the true positive rate
(*sensitivity*) and the false positive rate --- $1 -
\mbox{specificity}$ where *specificity* refers to the true negative
rate --- for varying probability thresholds according to the posterior
means of the predicted probabilities. For example, for a threshold of
approximately 0.1, the model will correctly predict 50 percent of
onsets while misclassifying only 8 percent of non-onsets. The total
performance of the model can be summarized by the area under the curve
(AUC), which in this case is `r auc`.[^auc] However, for the present
analysis the results from the ROC curve are misleading since there is
a significant class imbalance in the dataset. Onsets comprise only 4
percent of all observations meaning that the false positive rate is
less sensitive to changing thresholds. Figure \ref{fig:roc} therefore
also plots a precision-recall (PRC) curve, which depicts the tradeoff
between *precision* --- the number of true positives divided by the
sum of the true positives and false positives --- and recall --- the
number of true positives divided by the sum of the true positives and
false negatives --- for varying thresholds. The advantage is that
*precision* is invariant to the number of non-onset years in the
dataset. From this plot it is clear that the precision of the model,
the ability to discriminate between onset country-years and false
predicted non-set country-years, drops drastically as the *recall* of
the model increases. The *AUC-PR* for this plot is only `r auc_pr`.
[^auc_pr]

[^auc]: An AUC of 1 for a ROC curve indicates that the model is able
    to perfectly predict the outcome while an AUC of 0.5 denotes the
    performance of a purely random classifier.

[^auc_pr]: Similar to the ROC curve, an ACU of 1 for a PRC curve
    denotes a perfect model. The AUC for the null model, or random
    classifier, is the proportion of actual positive outcomes to negative
    outcomes in the dataset. In this case this means approximately
    .042.

```{r fig.width = 7, fig.cap = "\\label{fig:roc} hello"}

autoplot(x, curvetype = c("ROC", "PRC"))
```

Taken together, the evidence suggests a fairly poor fit to the
underlying data. The confusion matrix demonstrates that even with a
relatively low cutoff, the model is still only able to accurately
predict 25 percent of conflict onsets. According to the ROC curve,
decreasing this threshold improves the true positive rate with a
relatively decent tradeoff in terms of the false positive rate, but as
stated, this is largely due to the imbalance in the outcome
variable. In terms of *precision*, the performance of the model
quickly degrades as the threshold decreases.

### Discussion

- Link back to original theory, confirmatory

- Small effect size
- Poor fit
- Noisy Data
- Robustness?
- Compared to democracy?

Further analyses?
